% -*- LaTeX -*-
% -*- coding: utf-8 -*-
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%
%                             michael a.g. aïvázis
%                      california institute of technology
%                      (c) 1998-2010  all rights reserved
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%

\pagestyle{headandfoot}
\runningfootrule
\firstpageheader{ACM/CS 114}{Assignment 3}{Due: 8 Feb 2010}
\runningheader{}{}{}
\firstpagefooter{}{}{}
\runningfooter{ACM/CS 114}{Assignment 2}{\thepage}


% --------------------------------------
% Monte Carlo integration
\def\Li{\mbox{\rm Li}_{2}}
\def\dilog{\mbox{\tt dilog}}

An alternative expression for dilogarithm $\Li$ is
%
\begin{equation}
\Li(z) \bydef
- \int_{0}^{z} dz' \; \frac{\log(1-z')}{z'} \label{eq:li-double}
\end{equation}
%
which explains the name.

\begin{questions}

\question The effectiveness of computing using stochastic techniques is intimately related to
the properties of the pseudo-random generator used.

\begin{parts}
  \part Choose a random number generator and use it to fill a container with 10 numbers between
  0 and 1. Do you get the same 10 numbers every time you run the program? How can you seed your
  generator so that each execution of your program produces the same sequence? How can you seed
  it so that each execution produces different random numbers?

  \part Use your generator to fill a container with $10^6$ numbers between 0 and 1. Partition
  the interval $[0,1]$ into 10 bins of equal length and build a frequency histogram, i.e.~plot
  the number of times random numbers from your container fall within each bin. What is the
  average occupancy of the bins? What is the standard deviation?

  \part Repeat the above steps using 100 bins.

\end{parts}

\question Let's use Monte Carlo integration to approximate the value of the integral in
\eqref{li-double}. Implement the sequential version and test it for $z = 1$. How many integrand
evaluations are required to get the error to drop below $10^{-4}$ and $10^{-6}$?

\question Implement a parallel version using threads.

\question Implement a parallel version using \mpi.

\end{questions}


% end of file 
